Morse Theory

### 2020-12-01 paper.tex
added application

\begin{thebibliography}{9}
     \bibitem{s}
         {\sc }
         `` https://core.ac.uk/download/pdf/20670562.pdf
\end{thebibliography}
\end{document}

>>>>

\section{Applications}
\subsection{Topological Data Analysis}
A computational method for extracting simple descriptions of high dimensional data sets is a Mapper algorithm:
Let $X$ be a point cloud (data set) and we have a function $f:X \to \R$.
Assume that we have $N$ points in $X$ whose outputs are known for the function $f$.
Also, we assume that it is possible to compute inter-point dinstances between the points in the data.
Let $I$ be the range of the function $f$ restricted to the $N$ points and let $S$ be the covering of $I$.
We begin by dividing $S$ into a smaller intervals $(S_i)$ which overlap.
This gives us two parameters $l_i$ the length of $S_i$ and $p_{ij}$ the perceptange overlap between successive intervals.
Now for each $S_i$ we find a set $X_j = \{x | f(x) \in S_j\}$ s.t. $X_j \subset X$. The set $\{X_j\}$ forms a cover of $X$, and $X \subset \cup_j X_j$.
For each smaller set $X_j$ we find clusters $\{X_{jk}\} $.
We treat each cluster as a vertex in our reeb graph and draw an edge between vertices whenever $X_{jk} \cap X_{lm} \neq \emptyset$ i.e. the clusters corresponding to vertices have non-empty intersection. \\

\textbf{Ex.} Find a Reeb graph of a point cloud that has a shape of torus under the function $f(x,y,z) = z$. Approach the problem with assumption that we don't know the underlying distribution i.e. we don't know that it is a shape of torus. \\
Consider the following picture
\begin{center}
\includegraphics[width=1.0\textwidth]{mapper.png}
\end{center}
We have a map $f:N\to \R$ where N is a set of points in shape of torus.
We create a covering $S$ of the image $f(N)$ and we split the covering into 5 small overlaping intervals.
Now we consider $X_i = f^{-1}(S_i)$ and apply clustering algorithm on the output of the function.
For $X_1$, $X_2$, $X_4$ and $X_5$ there is only one cluster so we draw one vertex, but for $X_3$ we draw two.
Now we check that following clusters overlap: $X_{11} \cap X_{21} \neq \emptyset $, $X_{21} \cap X_{31} \neq \emptyset $, $X_{21} \cap X_{32} \neq \emptyset $, $X_{31} \cap X_{41} \neq \emptyset $,$X_{32} \cap X_{41} \neq \emptyset $,$X_{32} \cap X_{41} \neq \emptyset $.So there's a connection between vertices that represent those clusters. The output is a Reeb graph. \\ 
A famous example of application of Mapper algorithm is implemented by Nikolau, Levine and Carlsson in Topology based data analysis. They are able to identify a subgroup of breast cancers with a unique mutational profile and excellent survival.\\

We have implemented the Mapper algorithm and ran it for the sk learn diabetes dataset and standford cars dataset, output are as follows:

\begin{center}
\includegraphics[width=.4\textwidth]{diabetes.png}
\includegraphics[width=.4\textwidth]{cars.png}
\end{center}
Based on the diabetes data we may be able to make a difference between patients who don't need a treatmant, diabetis 1 and diabetis 2. While in the cars dataset it looks like there aren't any great differences between car features.
\begin{thebibliography}{9}
     \bibitem{s}
         {\sc }
         `` https://core.ac.uk/download/pdf/20670562.pdf
\end{thebibliography}
\end{document}
